Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass

### 2024-03-13 manuscript.Rmd
add supplementary figures and tables

The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure 3), as hypothesized (H2b). Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm). The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3). Overall, AGC stocks were a poor predictor of SOC stocks.

@@

```{r fig-distr-sol, fig.height = 5, fig.width = 15}
## diagnostic données sol pour marguerite
g1 <- ggplot(data_soil, aes(x = CF*100, col = SAMPLE_DEPTH)) +
  geom_histogram(bins = 30) +
  geom_text(data = subset(data_soil, CF > 0.3), angle = -60,
            aes(x = CF*100-5, y = 15,
                label = paste("Village:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
  labs(x = "Elements grossiers (% du volume)",
       y = "Nombre d'échantillons (site x profondeur)",
       col = "Profondeur (cm)") +
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()  +
  theme(legend.position = c(0.8, 0.9))

@@

ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
ggsave("figures/fig-distr-sol.pdf", height = 5, width = 15)
```

```{r fig-clay-corg, fig.height = 4, fig.width = 6}
ggplot(data_soil, aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
  geom_point() +
  labs(col = "Profondeur (cm)", x = "Argile (%)", y = "Teneur en carbone (%)") +
  geom_smooth(method = "lm") +
  theme_classic()
ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
```

@@

@@

```{r tetum-names}
sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
sheets2 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)

@@

vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
write.csv(vern, "data/species-correspondance.csv", row.names = FALSE)
```
# References

>>>>

The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure 3), as hypothesized (H2b). Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm). The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1)`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3). Overall, AGC stocks were a poor predictor of SOC stocks.

@@

```{r sfig-distr-sol, fig.height = 5, fig.width = 15}
## diagnostic données sol pour marguerite
g1 <- ggplot(data_soil, aes(x = CF*100, col = SAMPLE_DEPTH)) +
  geom_histogram(bins = 30) +
  geom_text(data = subset(data_soil, CF > 0.3), angle = -60,
            aes(x = CF*100-5, y = 15,
                label = paste("Village:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
  labs(x = "Elements grossiers (% du volume)",
       y = "Nombre d'échantillons (site x profondeur)",
       col = "Profondeur (cm)") +
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()  +
  theme(legend.position = c(0.8, 0.9))

@@

ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
ggsave("figures/fig-distr-sol.png", height = 5, width = 15)
```

```{r sfig-clay-corg, fig.height = 4, fig.width = 6}
data_soil[, village := substr(ID_VILLAGE, 1, 1)]
data_soil[village == "S", village := "Osso Luga"]
data_soil[village == "G", village := "Gariuai"]

ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
  geom_point() +
  labs(col = "Soil depth (cm)", x = "Clay content (%)", y = "SOC content (%)") +
  scale_color_brewer(type = "qual") +
  # geom_smooth(method = "lm") +
  theme_classic()
ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)

ggplot(data_soil[SAMPLE_DEPTH != "10-20"], aes(x = as.numeric(CLAY) + as.numeric(`FINE SILT`), y = Corg, col = SAMPLE_DEPTH, shape = village)) +
  geom_point() +
  labs(col = "Soil depth (cm)", x = "Clay and fine silt content (%)", y = "SOC content (%)", shape = "Site") +
  scale_color_brewer(type = "qual") +
  # geom_smooth(method = "lm") +
  theme_classic()
ggsave("figures/fig-clay-silt-corg.png", height = 4, width = 6)
```

@@

### Effect of altitude on SOC stocks

```{r sfig-altitude}
dataCelev = merge(dataC[, c("N_PLOT", "meanSC", "uprSC", "lwrSC")], 
                 data_plot[, c("N_PLOT", "elev", "village")], by = "N_PLOT")

ggplot(dataCelev, aes(x = elev, y = meanSC, col = village)) +
  geom_pointrange(aes(ymin=lwrSC, ymax = uprSC)) + 
  theme_classic() +
  labs(x = "Elevation (m)", y = "SOC stocks (Mg C/ha)", col = "Site")
ggsave("figures/soc-elev.png", height = 4, width = 6)
```

### Site soil characteristics

```{r stab-site-soil}
data_soil$silt <- as.numeric(data_soil$`FINE SILT`) + as.numeric(data_soil$`COARSE SILT`)
data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)

vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "CA", "Mg", "K", "Na", "CEC")
data_soil_plot = melt(data_soil, id.vars = c("village", "N_PLOT", "SAMPLE_DEPTH"), 
                      measure.vars = c("Corg\n", "N", "Corg_N", "CLAY", "silt", "sand", "pH\n", "Ca", "Mg\n", "K", "Na", "CEC\n"))

data_soil_plot[, variable := gsub("\n", "", variable)]

# when value below detection limit: use half of detection value
data_soil_plot[grepl("LD", value), value := as.numeric(strsplit(value, "=")[[1]][2])/2]
data_soil_plot[, value := as.numeric(value)]

# remove NA values
data_soil_plot = subset(data_soil_plot, !is.na(value))

# simulation 100 values per plot, soil depth and variable with a 10% coefficient
# of variation (CV value from lab)
data_soil_plot_simu = 
  data_soil_plot[, .(
    value = rnorm(100, value, sd = 0.1*value), n = 1:100), 
    .(village, N_PLOT, SAMPLE_DEPTH, variable)]

# bootstrap plot values to get mean value at the site level (for each simulation)
data_soil_site_simu = 
  data_soil_plot_simu[, .(
    value = mean(sample(value, length(value), replace = TRUE))),
    .(village, variable, n)]

# calculate mean and 95% confidence interval at site level
data_soil_site  = data_soil_site_simu[, .(
  value = paste0(
  signif(mean(value), 3), " (",
  signif(quantile(value, 0.025), 3), "-", 
  signif(quantile(value, 0.975), 3), ")"
  )), 
  .(village, variable)]

# change table shape to have one column per variable
data_soil_site = dcast(data_soil_site, village ~ variable, value.var = "value")
data_soil_site = data_soil_site[, c("village", "CLAY", "silt", "sand", "Corg", "N", "Corg_N", "pH", "Ca", "K", "Mg", "Na", "CEC")]
colnames(data_soil_site) = c("Site", "Clay (%)", "Silt (%)", "Sand (%)", "SOC content (%)","N", "C/N", "pH", "Ca", "K", "Mg", "Na", "CEC")

write.csv(data_soil_site, "data/soil_site.csv", row.names = FALSE)
```

@@

```{r tab-tetum-names}
sheets1 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_LARGE.xlsx"), value = TRUE)
sheets2 = grep("INV", readxl::excel_sheets("data/DATA_BIOMASS_TL_SHORT.xlsx"), value = TRUE)

@@

vern = vern[, .(latin = paste(latin, collapse = ", ")), .(tetum)]
write.csv(vern, "data/species-correspondance.csv", row.names = FALSE)
```

```{r timer-leste-c-footprint-lulcc}
df_eluc <- read_excel("C:/Users/piponiot-laroche/Downloads/National_LandUseChange_Carbon_Emissions_2023v1.0.xlsx", sheet = "OSCAR") 
colnames(df_eluc) <- df_eluc[7,]
df_eluc <- df_eluc[-(1:8),colnames(df_eluc) %in% c("unit: Tg C/year", "Timor-Leste")]
```

# References
